---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fable 
[![Travis build status](https://travis-ci.org/tidyverts/fable.svg?branch=master)](https://travis-ci.org/tidyverts/fable)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/tidyverts/fable?branch=master&svg=true)](https://ci.appveyor.com/project/tidyverts/fable)
[![Coverage status](https://codecov.io/gh/tidyverts/fable/branch/master/graph/badge.svg)](https://codecov.io/github/tidyverts/fable?branch=master)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

The R package *fable* provides methods and tools for displaying and analysing univariate time series forecasts including exponential smoothing via state space models and automatic ARIMA modelling. Data, model and forecast objects are all stored in a tidy format.

## Installation

You can install the **development** version from
[GitHub](https://github.com/tidyverts/fable)

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("tidyverts/fable")
```

## Example

```{r example, message=FALSE}
library(fable)
library(tsibbledata)
UKLungDeaths %>%
  model(ets = ETS(log(mdeaths))) %>%
  forecast
```

You can read more about the functionality of this package and the ideas behind it here: https://tidyverts.github.io/tidy-forecasting-principles/